Implicit assumptions and their impact on photometric redshift PDF performance in the context of LSST

### 2018-11-22 paper/tex/data.tex
add sentence with 2percen nondetect
add mention of blinding

To assign an SED to each galaxy, the {\it Adding Density Dependent Spectral Energy Distributions} (\textsc{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} procedure was used. This consisted of training an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED using a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited Sloan Digital Sky Survey Data Release 6 Value Added Galaxy Catalog \citep{Blanton:05}.  Each SDSS spectrum is fit with a sum of five SED components using the \textsc{k-correct \red{v4\_3?}} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}, thus each galaxy SED is parameterized as five weights for the basis SEDs. The distance to the spatial projected fifth-nearest neighbour was used as a proxy for local density in the SDSS training sample. For each simulated galaxy, a galaxy with similar absolute $r$-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy.  This process is done in such a way as to preserve the colour-density relation of galaxy environment.  Given the SED, absolute $r$-band magnitude and redshift, we computed apparent magnitudes in the six LSST filter passbands, $ugrizy$. We assigned magnitude errors in the six bands using the simple model described in \citet{Ivezic:08}, assuming full 10-year depth observations had been completed.  The number of total 30-second visits assumed when generating the photometric errors differs slightly from the fiducial numbers assumed for LSST: we assume 60 visits in u-band, 80 visits in g-band, 180 visits in r-band, 180 visits in i-band, 160 visits in z-band, and 160 visits in y-band.
%numbers taken from Alex Abate's photErrorModel.py script in PhotoZDC1 repository, which was used with nYrObs=10.
In the course of simulating Gaussian photometric errors, we add noise to objects fluxes, and some of these noisy fluxes will become negative in one or more bands.  We call such negative fluxes ``non-detections'' and signify them with a placholder magnitude of 99.0 in the catalog.  Thus, further mentions of ``non-detections'' refer to objects that would be ``looked at but not seen'' in multi-band forced photometry, and the photo-z codes will treat them as such.

\subsubsection{Selection of training and test sets}
\label{sec:buzztraining}
The total catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.  In order for statistical errors not to dominate, we need less than one million galaxies in our sample.  Several studies claim that only a few tens of thousands of spectra are necessary to calibrate photo-z surveys to Stage IV requirements (e.~g.~\citet{Bernstein:10}, \citet{Masters:2017}).  Therefore, we aim for a final number of training galaxies between $3\times 10^{4}$ and $5\times 10^{4}$ in our sample.
%This catalogue contained two orders of magnitude more galaxies than were necessary to determine statistics for this study,
In order to reduce our sample to a reasonable size, we limit our dataset to a subset of $\sim\!16.8$ square degrees selected from five separate spatial regions of the simulation. Systematic problems with galaxy colors above $z\!>\!2$ were observed, so the catalogue was limited to include only galaxies in the redshift range $0\!<\!z\!\leq\!2.0$. A random subset of the the remaining galaxies was chosen, and placed at random into either a ``training'' set ($10$ per cent of the sample), for which the galaxies true redshifts will be supplied, or a ``test'' set (the remaining $90$ per cent of the sample), for which each code will need to predict a redshift PDF for each galaxy. %The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
Finally, we restrict our analysis to a sample with an apparent magnitude limite $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.  This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.  All subsequent results will evaluate this ``gold sample'' test set.

>>>>

To assign an SED to each galaxy, the {\it Adding Density Dependent Spectral Energy Distributions} (\textsc{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} procedure was used. This consisted of training an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED using a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited Sloan Digital Sky Survey Data Release 6 Value Added Galaxy Catalog \citep{Blanton:05}.  Each SDSS spectrum is fit with a sum of five SED components using the \textsc{k-correct \red{v4\_3?}} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}, thus each galaxy SED is parameterized as five weights for the basis SEDs. The distance to the spatial projected fifth-nearest neighbour was used as a proxy for local density in the SDSS training sample. For each simulated galaxy, a galaxy with similar absolute $r$-band magnitude and local galaxy density was chosen from the training set, and that training galaxy's SED was assigned to the simulated galaxy.  This process is done in such a way as to preserve the colour-density relation of galaxy environment.  Given the SED, absolute $r$-band magnitude and redshift, we computed apparent magnitudes in the six LSST filter passbands, $ugrizy$. We assigned magnitude errors in the six bands using the simple model described in \citet{Ivezic:08}, assuming full 10-year depth observations had been completed.  The number of total 30-second visits assumed when generating the photometric errors differs slightly from the fiducial numbers assumed for LSST: we assume 60 visits in u-band, 80 visits in g-band, 180 visits in r-band, 180 visits in i-band, 160 visits in z-band, and 160 visits in y-band.
%numbers taken from Alex Abate's photErrorModel.py script in PhotoZDC1 repository, which was used with nYrObs=10.
In the course of simulating Gaussian photometric errors, we add noise to objects fluxes, and some of these noisy fluxes will become negative in one or more bands.  We call such negative fluxes ``non-detections'' and signify them with a placholder magnitude of 99.0 in the catalog.  Thus, further mentions of ``non-detections'' refer to objects that would be ``looked at but not seen'' in multi-band forced photometry, and the photo-z codes will treat them as such.  Only 2.0 per cent of our sample galaxies contain a photometric band with a non-detection, the vast majority of which are in the $u$-band.

\subsubsection{Selection of training and test sets}
\label{sec:buzztraining}
The total catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.  In order for statistical errors not to dominate, we need less than one million galaxies in our sample.  Several studies claim that only a few tens of thousands of spectra are necessary to calibrate photo-z surveys to Stage IV requirements (e.~g.~\citet{Bernstein:10}, \citet{Masters:2017}).  Therefore, we aim for a final number of training galaxies between $3\times 10^{4}$ and $5\times 10^{4}$ in our sample.
%This catalogue contained two orders of magnitude more galaxies than were necessary to determine statistics for this study,
In order to reduce our sample to a reasonable size, we limit our dataset to a subset of $\sim\!16.8$ square degrees selected from five separate spatial regions of the simulation. Systematic problems with galaxy colors above $z\!>\!2$ were observed, so the catalogue was limited to include only galaxies in the redshift range $0\!<\!z\!\leq\!2.0$. A random subset of the the remaining galaxies was chosen, and placed at random into either a ``training'' set ($10$ per cent of the sample), for which the galaxies true redshifts will be supplied, or a ``test'' set (the remaining $90$ per cent of the sample), for which each code will need to predict a redshift PDF for each galaxy.  
%The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
Finally, we restrict our analysis to a sample with an apparent magnitude limite $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.  This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.  All subsequent results will evaluate this ``gold sample'' test set.  In order to blind results, initially redshifts were not revealed for the ``test'' set, and were only supplied for the training sample galaxies.  This prevented code runners from tweaking results and over-fitting to the specific test set.
